%%% Die folgende Zeile nicht ändern!
\section*{\ifthenelse{\equal{\sprache}{deutsch}}{Zusammenfassung}{Abstract}}
%%% Zusammenfassung:

This bachelor thesis extensively evaluates the vulnerability of capsule network to various adversarial attacks.
Capsule networks are a class of artificial neural network in which neurons are grouped in capsules whose activity vectors describe different properties of the same entity.
Recent work attributes them more robustness to adversarial attacks than other types of neural networks.

To investigate those conjectures we train four pairs of capsule networks and convolutional neural networks to classify different image datasets.
Using multiple strong and diverse adversarial attack we compute adversarial examples for each combination of dataset and architecture.

By comparing the Euclidean norms of the generated perturbations we find that capsule networks are vulnerable to adversarial attacks in a similar way as convolutional networks.
We discover in particular that capsule networks show no robustness to sufficiently strong white-box attacks.
Additionally, we demonstrate limited transferability of the adversarial examples between the two types of architectures.

Furthermore, we study the structure of the adversarial perturbations and observe that they seem to lie in a relatively low dimensional linear subspace for both types of networks.
Despite those similarities we detect inherent differences between the perturbations constructed for the different network types using dimensionality reduction.

Finally, we aim to identify the cause of adversarial vulnerability in the networks we trained.
Since capsule networks and convolutional networks partially use the same type of convolutional layer, we focus on the disturbance on the activation of individual layers caused by adversarial attacks.
We find that susceptibility to attacks is not just ascribable to the first layers in a network but a result of deeper layers as well.
Specifically, we find that the dynamic routing algorithm used in capsule networks does not establish capsule activation invariance to counteract adversarial attacks as previously conjectured.
